HEP Software Foundation Community White Paper Working Group -- Detector Simulation

### 2018-02-21 CWP/papers/HSF-CWP-2017-07_detsim/latex/DetectorSimulation.tex
added text and reference on LiAr TPCs in section on Intensity Frontier

Major components of the intensity frontier over the next decade will
include liquid Argon TPC detectors and liquid Xenon calorimeters.
Efforts to address common needs of the liquid Argon experiments
including the post Geant4 phase of simulation are undertaken in projects
such as LArSoft~\cite{LArSoft} and Qscan~\cite{LArLEM-TPC}, while extending Geant4 EM
physics models for liquid noble gases is undertaken by the NEST project
~\cite{Szydagis:2013sih}. Detectors with large numbers of wires and photo-sensors may
naturally lend themselves to parallelisation techniques, provided that
the simulation frameworks are able to handle such cases in an efficient
way. Additionally, the Wire Cell Toolkit ~\cite{Wire-Cell} provides simulation,
noise filtering, signal processing and reconstruction algorithms for
LArTPC detectors. It includes experimental support for parallel
processing following the data flow processing paradigm. More work and
input from parallel processing experts is needed to bring this support
to maturity.

>>>>

Major components of the intensity frontier over the next decade will
include liquid Argon TPC detectors and liquid Xenon calorimeters.
For example, liquid Argon TPC's are equipped with optical readout systems 
to detect optical photons from the scintillation process. The information 
from the optical system is used to  provide a trigger and, potentially, 
could be used to improve the energy resolution of showers by combining the 
scintillation and charge collection signals. The modeling of optical photons 
in liquid Argon detectors is a real challenge, since 50,000 scintillation 
photons are produced per 1 MeV of energy deposited in liquid Argon. To track 
all these photons in Geant4 would increase the computing time on a CPU by a 
factor of  more than one thousand. Therefore, the full simulation is only used 
to produce lookup tables with the probability that that an optical photon makes 
it to the photo detector. These tables are an approximation and cannot replace 
the full simulation without a significant loss in physics accuracy. Even the 
lookup table solution requires a significant amount of CPU time to produce 
enough statistics. In order to satisfy the physics accuracy and computing 
performance requirements of future neutrino experiments, approaches based 
on new computing platforms are under investigation. For example, 
Opticks ~\cite{1742-6596-898-4-042001} 
is a software package where the optical physics processes are implemented on the GPU. 
This package is executed on NVIDIA machines to parallelize the
tracing of optical photons and yields a drastic time performance improvement 
with a speedup of the order of a thousand with respect to the original Geant4 
package running on CPUs. Opticks has been used to simulate the optical photons 
in the JUNO neutrino experiment.

Efforts to address common needs of the liquid Argon experiments
including the post Geant4 phase of simulation are undertaken in projects
such as LArSoft~\cite{LArSoft} and Qscan~\cite{LArLEM-TPC}, while extending Geant4 EM
physics models for liquid noble gases is undertaken by the NEST project
~\cite{Szydagis:2013sih}. Detectors with large numbers of wires and photo-sensors may
naturally lend themselves to parallelisation techniques, provided that
the simulation frameworks are able to handle such cases in an efficient
way. Additionally, the Wire Cell Toolkit ~\cite{Wire-Cell} provides simulation,
noise filtering, signal processing and reconstruction algorithms for
LArTPC detectors. It includes experimental support for parallel
processing following the data flow processing paradigm. More work and
input from parallel processing experts is needed to bring this support
to maturity.
